Impact on End-users by ISP IPv6 Deployment

### 2022-11-22 paper.tex
draft 1.0 finished todo: check spelling and sentence structure

    This last section deals with ISP tunnels. Especially at the beginning of IPv6 deployment, IPv6 path stablility was worse than IPv4 path stablility\cite{7119767}. That is because of the way ISPs transport IPv6 traffic. This has an impact on end-users. ISPs could build an IPv4 only and an IPv6 only infrastructure. To make the transition faster, ISPs tunnel IPv6 traffic over existing IPv4 paths. There are many protocols to achieve that. Two of them will be discussed now.

@@

    \section{Summary}
    With the exhaustion of IPv4 addresses, ISPs are forced to transition to IPv6 addresses. This has an direct impact on the end user, becasue both protocols are not compatible. Before transitioning to IPv6, ISPs invested heavily in NAT systems. This caused customer end-devices to not be routable in the public internet. NAT effected many protocols. NAT traversal

>>>>

    This last section deals with ISP tunnels. Especially at the beginning of IPv6 deployment, IPv6 path stablility was worse than IPv4 path stablility\cite{7119767}. That is because of the way ISPs transport IPv6 traffic. This has an impact on end-users. ISPs could build an IPv4 only and an IPv6 only infrastructure. To make the transition faster, ISPs tunnel IPv6 traffic over existing IPv4 paths. There are many protocols to achieve that. It is also possible to tunnel IPv4 over IPv6. Existing IPv4 dominance does not hinder IPv6 only investments because of that. In the following two protocols of IPv6 in IPv4 tunneling will be discussed.

@@

    \section{Summary}
    With the exhaustion of IPv4 addresses, ISPs are now forced to transition to IPv6 addresses. This has an direct impact on the end user, becasue both protocols are not compatible.\\\\Before transitioning to IPv6, ISPs invested heavily in NAT systems. This caused customer end-devices to not be routable in the public internet. \textit{NAT44} effected many protocols and required a new version of them. NAT traversal can be a security problem. This is why NAT should be abandoned as soon as possible. With \textit{NAT444} even more problems appeared for the end-user.\\\\\textit{Dualstack} introduced IPv6 addresses to end-users for the first time. This enabled user to host multiple public, globally routable servers from their home. With a further scarcity of IPv4 addresses, \textit{dualstack lite} was introduced. \textit{Dualstack lite} outsources the NAT from the end-user to ISP level. The end-user can only profit from a \textit{dualstack} connection. There are only added features.\\\\Today, ISPs start to provide IPv6 only services. With \textit{NAT64/DNS64} end-users are only provided one IPv6 block. Conncection to IPv4 only servers is done over ISP site translation. \textit{464XLAT} performs stateless translation at end-user level and statefull translation at ISP level. Both systems allow the ISP too share precious public IPv4 addresses amoung multiple customers. In most cases these IPv4 addresses are not needed, because the end-users use a nativ IPv6 connection to the servers they want to access. Directly compared to \textit{dualstack}, IPv6 only solutoins are worse for the end-user. Translation is slower than a direct connection and some features of IPv4 do not work. It is not possible to reach an IPv6 only connected home over IPv4.\\\\With IPv6 addresses for every device, applications running on the users end-device need to support IPv6 too. \textit{Bump-In-the-Host} is a system, to use IPv6 even with legacy application that only support IPv4. Making old applications work with IPv6 has only an positiv impact for the end-user\\\\To improve IPv6 availability, ISPs tunnel IPv6 traffic over existing IPv4 infrastructure. This has a positiv effect on the path availability and thus a positiv effect on the end-user experience. \textit{6in4} is a very basic protocol to establish those tunnels. \textit{6rd} is an improved version. The end-user profits from that.\\\\All in all, the deployment of IPv6 by ISPs benefits the end-user, at least in the long term. There might be some devices or applications not work properly at the beginning. Though, once the transition is completed, end-users have great possibilities. IPv6 enables all devices to initiate true end-to-end commuinication. End-users are able to host multiple servers and participate in peer-to-peer networks. But even in the short term, users profit from IPv6. For example, online multiplayer games work better with IPv4, because no NAT is involved. To answer the title of this paper in one sentence:\\\\The end-user impact is positiv.
